---
title: "07-last steps SE"
author: "Jelmer"
date: "25-1-2021"
output: html_document
---

#### Last steps before SummarizedExperiment

Now we can load the CSV_file, you’ll see the first few counts inside the file. Check if the data file is a “data.frame.

```{r}
#Load the raw data from the csv data
load_csv_data <- list.files(path = here::here(params$geo_dataset), full.names = TRUE, pattern = ".csv")
csv_data <- read.csv(load_csv_data)
#the first few counts of the raw data
head(csv_data)

#Check if the csv_data is a data.frame
class(csv_data)
```

In raw_counts_csv_data we load in the raw counts of the csv file, minus the first column because these are names of the samples. Also note that it can be more or less columns in different documents. In the next step we change the name of the columns to make them more understandable. The last step is to change the data.frame into a data.matrix for SummarizedExperiment.
```{r}
## In het volgende stukje willen we dat de namen goed komen te staan, je ziet eerst dat het rommelig is en daarna hebben ze GSM nummers
#Laat alleen de raw counts over
raw_counts_csv_data <- csv_data[2:21]
#head(raw_counts_csv_data)

##make rowdata_csv
rowdata_csv_data <- raw_counts_csv_data$`Gene symbol`

#Kan ik de raw_counts_csv_data gebruiken om colnames te maken test
colnames(raw_counts_csv_data) <-samplenames_csv[[1]]
head(raw_counts_csv_data)

#maak er een matrix van, weet niet waarom

raw_counts_csv_data_matrix <- data.matrix(raw_counts_csv_data)
rownames(raw_counts_csv_data_matrix) <- rowdata_csv_data
#head(raw_counts_csv_data_matrix)
```

SummarizedExperiment is used for organising data. This package saves rectangular matrices from experimental results. The rows represent the characteristics of interest, typically genes and transcripts. The columns stand for the samples. ColData consits of samples and metadata. Metadata is data about data, so ColData contains extra information about the samples. 

Here we will put a character factor in a word. The word now consits of the organised data of the raw sample data coming from the dataset. 
```{r}
se_csv_data <- SummarizedExperiment(assays = raw_counts_csv_data_matrix,
                     rowData = rowdata_csv_data,
                     colData = phenodata_csv)
```

Here we will ad the metadata manually because the SummarizedExperiment won't accept it otherwise. The $ function extracts elements from a named list, by heading. 
```{r}
metadata(se_csv_data)$metadata <- metadata 
```

With the code below we will be able to inspect the SummarizedExperiment. The code "head" will make it possible to view the first part of the data frame. 
```{r}
se_csv_data
head(assay(se_csv_data))
```


#### iSEE

This function can be used to explore the data stored in the SummarizedExperiment, like the row- and column-level metadata. 
```{r}
#iSEE test
#se_csv_iSEE <- se_csv_data
#iSEE(se_csv_iSEE)
```
